---
title: "Group 1 Final"
author: "Evelyn Campo, Xiao Qi, Nusrat Prithee, Roman Kosarzycki"
date: "today"
# date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r init, include=FALSE}
library(ezids)
library(ggplot2)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
```


```{r}

heartdata <- data.frame(read.csv("heart_2020_cleaned.csv"))

n=10
#heartdata <- heartdata[seq(1,nrow(heartdata),n),]

heartdata$HeartDisease <- as.factor(heartdata$HeartDisease)
heartdata$Smoking <- as.factor(heartdata$Smoking)
heartdata$AlcoholDrinking <- as.factor(heartdata$AlcoholDrinking)
heartdata$Stroke <- as.factor(heartdata$Stroke)
heartdata$DiffWalking <- as.factor(heartdata$DiffWalking)
heartdata$Sex <- as.factor(heartdata$Sex)
heartdata$Race <- as.factor(heartdata$Race)
heartdata$Diabetic <- as.factor(heartdata$Diabetic)
heartdata$PhysicalActivity <- as.factor(heartdata$PhysicalActivity)
heartdata$GenHealth <- as.factor(heartdata$GenHealth)
heartdata$Asthma <- as.factor(heartdata$Asthma)
heartdata$KidneyDisease <- as.factor(heartdata$KidneyDisease)
heartdata$SkinCancer <- as.factor(heartdata$SkinCancer)

heartdata$AgeCategory <- gsub(" or older","-4",as.character(heartdata$AgeCategory))
heartdata$LoAge <- as.numeric(substr(heartdata$AgeCategory,1,2))
heartdata$rand <- sample(0:4,size=nrow(heartdata),replace=T)
heartdata$Age <- with(heartdata,LoAge+rand)
heartdata <- subset(heartdata,select=-c(AgeCategory,LoAge,rand))

str(heartdata)

```
### Testing

###Chi-Square Test

**What variables affect mental health physical health? In particular, does alcohol drinking, smoking**

**Does the Heat Disease effect on Gen Health??**
```{r, results="markup"}
contable1= table(heartdata$HeartDisease,heartdata$GenHealth)
ezids::xkabledply(contable1, title="Contingency table for Heart Disease  vs Gen Health ")
chitest1 = chisq.test(contable1)
chitest1
```
**Is heart disease related to gender??**

```{r, results="markup"}
contable2= table(heartdata$HeartDisease,heartdata$Sex)

ezids::xkabledply(contable2, title="Contingency table for Heart Disease  vs Gender ")
chitest1 = chisq.test(contable2)
chitest1

```

**What is the average age people are suffering from heart disease??**
```{r, results="markup"}
 heart_disease_on=subset(heartdata,HeartDisease=="Yes")
ttest_age <- t.test(heart_disease_on$Age)
ttest_age
```

```{r, results="markup" }
library(ggplot2)
ggplot(data=heart_disease_on,aes(x=Age,)) + 
geom_histogram(col="red", 
               fill="green", 
               alpha = .7) + 
labs(title="Age of suffering histogram") +
labs(x="Age", y="frequency")
```
**What variables affect mental health physical health? In particular, does alcohol drinking, smoking**


