---
title: "Group 1 Final"
author: "Evelyn Campo, Xiao Qi, Nusrat Prithee, Roman Kosarzycki"
date: "today"
# date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r init, include=FALSE}
library(ezids)
library(ggplot2)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
```


```{r}

heartdata <- data.frame(read.csv("heart_2020_cleaned.csv"))

n=10
#heartdata <- heartdata[seq(1,nrow(heartdata),n),]

heartdata$HeartDisease <- as.factor(heartdata$HeartDisease)
heartdata$Smoking <- as.factor(heartdata$Smoking)
heartdata$AlcoholDrinking <- as.factor(heartdata$AlcoholDrinking)
heartdata$Stroke <- as.factor(heartdata$Stroke)
heartdata$DiffWalking <- as.factor(heartdata$DiffWalking)
heartdata$Sex <- as.factor(heartdata$Sex)
heartdata$Race <- as.factor(heartdata$Race)
heartdata$Diabetic <- as.factor(heartdata$Diabetic)
heartdata$PhysicalActivity <- as.factor(heartdata$PhysicalActivity)
heartdata$GenHealth <- as.factor(heartdata$GenHealth)
heartdata$Asthma <- as.factor(heartdata$Asthma)
heartdata$KidneyDisease <- as.factor(heartdata$KidneyDisease)
heartdata$SkinCancer <- as.factor(heartdata$SkinCancer)

heartdata$AgeCategory <- gsub(" or older","-4",as.character(heartdata$AgeCategory))
heartdata$LoAge <- as.numeric(substr(heartdata$AgeCategory,1,2))
heartdata$rand <- sample(0:4,size=nrow(heartdata),replace=T)
heartdata$Age <- with(heartdata,LoAge+rand)
heartdata <- subset(heartdata,select=-c(AgeCategory,LoAge,rand))

str(heartdata)

```
### Testing

###Chi-Square Test

**Does the data support that race very much affects heart disease???**

```{r, results="markup"}
racetable = xtabs(~ Race+HeartDisease, data =heartdata )
racetable
chitest_race = chisq.test(racetable)
chitest_race
```

**Does the data support that Heat Disease effect on Gen Health??**
```{r, results="markup"}
gentable= table(heartdata$HeartDisease,heartdata$GenHealth)
ezids::xkabledply(gentable, title="Contingency table for Heart Disease  vs Gen Health ")
chitest_gen = chisq.test(gentable)
chitest_gen
```
**Is heart disease data related to gender??**

```{r, results="markup"}
sextable= table(heartdata$HeartDisease,heartdata$Sex)

ezids::xkabledply(sextable, title="Contingency table for Heart Disease  vs Gender ")
chitest_sex = chisq.test(sextable)
chitest_sex

```
###T-test
**What is the average age people are suffering from heart disease??**
```{r, results="markup"}
 heart_disease_on=subset(heartdata,HeartDisease=="Yes")
str(heart_disease_on)
ttest_age <- t.test(heart_disease_on$Age)
ttest_age
```

###Test For Association

**What variables affect mental health physical health? In particular, does alcohol drinking, smoking**
```{r results='markup'}
menhealth_smoke<-cor.test(heartdata$MentalHealth,as.numeric(heartdata$Smoking), method="pearson")
menhealth_smoke
menhealth_drinking<-cor.test(heartdata$MentalHealth,as.numeric(heartdata$AlcoholDrinking), method="pearson")
menhealth_drinking
```  

```{r results='markup'}
phyhealth_smoke<-cor.test(heartdata$PhysicalHealth,as.numeric(heartdata$Smoking), method="pearson")
phyhealth_smoke
phyhealth_drinking<-cor.test(heartdata$PhysicalHealth,as.numeric(heartdata$AlcoholDrinking), method="pearson")
phyhealth_drinking
```  