---
title: "Group 1 Midterm"
author: "Evelyn Campo, Xiao Qi, Nusrat Prithee, Roman Kosarzycki"
date: "March 9, 2022"
# date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r init, include=FALSE}
library(ezids)
library(ggplot2)
library(dplyr)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
```

Importing the Dataset assigning "NA" to all blank cells
```{r}
tvdata <- read.csv("tv_shows.csv", header=T, na.strings=c("","NA")) 
```


Cleaning the Dataset
```{r}
tvdata$IMDb<-gsub("/10","",as.character(tvdata$IMDb)) #We had to delete the character /10 on every observation 
tvdata$Rotten.Tomatoes<-gsub("/100","",as.character(tvdata$Rotten.Tomatoes)) #We had to delete the character /100 on every observation 
tvdata$IMDb<-as.double(tvdata$IMDb) #as.double is a generic function. It is identical to as. numeric
tvdata$Rotten.Tomatoes<-as.double(tvdata$Rotten.Tomatoes) #as.double is a generic function. It is identical to as. numeric
tvdata = select(tvdata, -X, -ID, -Type) #This variables aren't of our interest

str(tvdata)
# View(tvdata)
```

Counting missing values 
```{r}
sum(is.na(tvdata)) #Total NA
mean(is.na(tvdata))

#NA per variables
sum(is.na(tvdata$Title))
sum(is.na(tvdata$Year))
sum(is.na(tvdata$Age))
sum(is.na(tvdata$IMDb))
sum(is.na(tvdata$`Rotten Tomatoes`))
sum(is.na(tvdata$Netflix))
sum(is.na(tvdata$Hulu))
sum(is.na(tvdata$`Prime Video`))
sum(is.na(tvdata$`Disney+`))

```
Age has the most missing values in the Dataset, and IMDb is the other variable with NA values.

Streaming platforms into as.factor()
```{r}
tvdata$Netflix <- as.factor(tvdata$Netflix)  
tvdata$Hulu <- as.factor(tvdata$Hulu)  
tvdata$Prime.Video <- as.factor(tvdata$Prime.Video)
tvdata$Disney. <- as.factor(tvdata$Disney.)

```
The streaming platforms are categorical. 1 represents if the TV show is found on their platform otherwise 0.


Just undertanding a lilttle more about streaming platform variables and age
```{r}

print("Age:")
table(tvdata$Age)
print("Netflix:")
table(tvdata$Netflix)
print("Hulu:")
table(tvdata$Hulu)
print("Prime Video:")
table(tvdata$Prime.Video)
print("Disney:")
table(tvdata$Disney.)

```


What are the most targeted age groups for the TV shows by Netflix, Hulu, Prime, disney
Video? 

```{r}

print("Age:")
table(tvdata$Age)
prop.table(table(tvdata$Age))

ggplot(data = na.omit(tvdata), aes(Age)) + 
  geom_bar(col="red", 
                 fill="blue", 
                 alpha = .7) + # opacity
  labs(title="Most targeted age group for all platforms: 16+ ") +
  labs(x="Target age of viewers ", y="Frequency")

#I am taking this graph from above since it works for this question
```
people who are 16 and older are the most targeted age groups for the tv shows among the all steamming platfrom.

.Which year published the highest number of TV shows? 
```{r}
print("Year:")
table(tvdata$Year)
library(ggplot2)
ggplot(data=tvdata, aes(Year)) + 
  geom_histogram(bins = 30,
                 col="red", 
           fill="blue", 
           alpha = .7) + # opacity
  labs(title="Year during which the highest number of shows were published: 2017") +
  labs(x="Year tv show was published", y="Frequency")
#I am replacing the prevous code with the graphs from above to keep the same formatting
```
The highest number of TV shows were published in 2017 (685) and 2018 (562).And the histogram is right skewed which indicates that video publication is raising while time goes forward.

Normality check of Variables

```{r}

netflixtv <- subset(tvdata, Netflix==1)
hulutv <- subset(tvdata, Hulu==1)
primetv <- subset(tvdata, Prime.Video==1)
disneytv <- subset(tvdata, Disney.==1)

```

Which streaming platform has the highest average rating (according to Rotten 
Tomatoes and IMDb)? 

```{r q06}
net_rating<-t.test(netflixtv$IMDb,netflixtv$Rotten.Tomatoes)
net_rating
hulu_rating<-t.test(hulutv$IMDb,hulutv$Rotten.Tomatoes)
hulu_rating
prime_rating<-t.test(primetv$IMDb,primetv$Rotten.Tomatoes)
prime_rating
disney_rating<-t.test(disneytv$IMDb,disneytv$Rotten.Tomatoes)
disney_rating
```
 Prime Videos has the highest average Imdb rating  which is 7.152538. Also, Netflix has the highest average RTT rating  which is 53.559107
 
Normality check for the variable IMDb and Rotten.Tomatoes
```{r}
shapiro.test(netflixtv$IMDb)
shapiro.test(netflixtv$Rotten.Tomatoes)

library(ggplot2)
ggplot(data=netflixtv,aes(x=IMDb,)) + 
geom_histogram(col="red", 
               fill="green", 
               alpha = .7) + 
labs(title="Netflix IMDb histogram") +
labs(x="IMDB", y="frequency")

ggplot(data=netflixtv,aes(x=Rotten.Tomatoes,)) + 
geom_histogram(col="red", 
               fill="green", 
               alpha = .7) + 
labs(title="Netflix Rotten.Tomatoes histogram") +
labs(x="Rotten.Tomatoes", y="frequency")


```
Here p value for IMDb and Rotten.Tomatoes is less than 0.05 for Netflix. That's mean these are not normally distributed. Histogram of IMDb is right-skewed also Histogram of Rotten.Tomatoes is slightly left-skewed.
```{r}
shapiro.test(hulutv$IMDb)
shapiro.test(hulutv$Rotten.Tomatoes)

library(ggplot2)
ggplot(data=hulutv,aes(x=IMDb,)) + 
geom_histogram(col="red", 
               fill="green", 
               alpha = .7) + 
labs(title="Hulu IMDb histogram") +
labs(x="IMDB", y="frequency")

ggplot(data=hulutv,aes(x=Rotten.Tomatoes,)) + 
geom_histogram(col="red", 
               fill="green", 
               alpha = .7) + 
labs(title="hulu Rotten.Tomatoes histogram") +
labs(x="Rotten.Tomatoes", y="frequency")

```
Here p value for IMDb and Rotten.Tomatoes is less than 0.05 for Hulu. That's mean these are not normally distributed. Histogram of IMDb is right-skewed also Histogram of Rotten.Tomatoes is slightly Bimodal.
```{r}
shapiro.test(primetv$IMDb)
shapiro.test(primetv$Rotten.Tomatoes)
library(ggplot2)
ggplot(data=primetv,aes(x=IMDb,)) + 
geom_histogram(col="red", 
               fill="green", 
               alpha = .7) + 
labs(title="prime IMDb histogram") +
labs(x="IMDB", y="frequency")

ggplot(data=primetv,aes(x=Rotten.Tomatoes,)) + 
geom_histogram(col="red", 
               fill="green", 
               alpha = .7) + 
labs(title="prime Rotten.Tomatoes histogram") +
labs(x="Rotten.Tomatoes", y="frequency")

```
Here p value for IMDb and Rotten.Tomatoes is less than 0.05 for Prime Videos. That's mean these are not normally distributed. Histogram of IMDb is right-skewed also Histogram of Rotten.Tomatoes is Randomly distribution.

```{r}
shapiro.test(disneytv$IMDb)
shapiro.test(disneytv$Rotten.Tomatoes)
library(ggplot2)
ggplot(data=disneytv,aes(x=IMDb,)) + 
geom_histogram(col="red", 
               fill="green", 
               alpha = .7) + 
labs(title="disney IMDb histogram") +
labs(x="IMDB", y="frequency")

ggplot(data=disneytv,aes(x=Rotten.Tomatoes,)) + 
geom_histogram(col="red", 
               fill="green", 
               alpha = .7) + 
labs(title="disney Rotten.Tomatoes histogram") +
labs(x="Rotten.Tomatoes", y="frequency")

```
Here p value for IMDb and Rotten.Tomatoes is less than 0.05 for Disney+. That's mean these are not normally distributed. Histogram of IMDb is slightly right-skewed and randomly distributed. also Histogram of Rotten.Tomatoes is slightly left-skewed and randomly distribution.

----------------

# SMART Questions

After an initial examination of our chosen data set, we decided on three SMART questions to focus on:

What are the most targeted age groups for the TV shows by Netflix, Hulu, PrimeVideo, Disney+?

Which year published the highest number of TV shows?

Which streaming platform has the highest average rating (according to Rotten Tomatoes and IMDb)?

The first question focuses on the relation between the target age group and the platform. By looking at the distribution of the column Age for each platform, we can obtain knowledge about the intended audience for Netflix, Hulu, Prime, and Disney. The second question focuses on the column Year. We quickly noticed that the range of years listed was larger than expected, spanning from 1904 to 2021. We expected that more recent years would have more listed TV shows, but we wanted to explore that distribution in more detail. Finally, the third question looks at the IMDb and Rotten Tomatoes ratings for each platform. While there were some TV shows that were available on more than one platform, we were interested in seeing how the overall ratings were distributed for each platform. 

During our exploratory data analysis, we also came up with a fourth SMART question:

What is the relationship between IMDb and Rotten Tomatoes?

We had not initially considered the idea that the IMDb and Rotten Tomatoes rating systems would differ by much, but more in-depth analysis revealed that there were significant differences between how the two systems rated shows. This created an additional dimension of analysis that we did not initially anticipate.

# Exploratory Data Analysis

## Rating Systems

### Frequency of Ratings

First, we wanted to use exploratory data analysis (EDA) to learn more about the two rating systems: IMDb and Rotten Tomatoes.

```{r results='markup'}

ggplot(tvdata, aes(x=IMDb)) + geom_histogram(binwidth = 0.1)+ggtitle("Frequency of IMDb Ratings") +geom_vline(xintercept = mean(tvdata$IMDb,na.rm = TRUE), color = "red", size=1.5)+geom_vline(xintercept = median(tvdata$IMDb,na.rm = TRUE), color = "blue", size=1.5)+geom_vline(xintercept = 7.4, color = "orange", size=1.5)

ggplot(tvdata, aes(x=Rotten.Tomatoes)) + geom_histogram(binwidth = 1)+ggtitle("Frequency of Rotten Tomatoes Ratings")+geom_vline(xintercept = mean(tvdata$Rotten.Tomatoes,na.rm = TRUE), color = "red", size=1.5)+geom_vline(xintercept = median(tvdata$Rotten.Tomatoes,na.rm = TRUE), color = "blue", size=1.5)+geom_vline(xintercept = 10, color = "orange", size=1.5)

# table(tvdata$IMDb)
# table(tvdata$Rotten.Tomatoes)

print("Mean IMDb rating:")
mean(tvdata$IMDb,na.rm = TRUE)
print("Median IMDb rating:")
median(tvdata$IMDb,na.rm = TRUE)
print("Mode IMDb rating:")
print(7.4)

print("Mean RT rating:")
mean(tvdata$Rotten.Tomatoes,na.rm = TRUE)
print("Median RT rating:")
median(tvdata$Rotten.Tomatoes,na.rm = TRUE)
print("Mode RT rating:")
print(10)

```

IMDb is a rating scale from 1 to 10, but the lowest rating on this list is 1.1 and the highest is 9.6. Rotten Tomatoes rates TV shows on a scale from 1-100 with a lowest score of 10 and a highest score of 100. As seen in the histogram plot, the distribution of IMDb ratings has a slight left skew. This is further exemplified by the fact that the median (blue) is larger than the mean (red). The Rotten Tomatoes ratings mean and median are almost equal, but there are some outliers in the data at the lower end of the rating system. The mode is a 10/100 with 304 shows receiving that rating. The Rotten Tomatoes rating is a combination of critics' ratings and audience ratings, but this data set only shows the total rating, which is a limitation of this dataset. It would have been interesting to see how critics and the audience agree or disagree about certain ratings. 
In comparing these ratings distribution, it became obvious that IMDb, on average, gives higher ratings than Rotten Tomatoes. IMDb has a mean of 7.09/10 (70.9%) and a median of 7.3/10 (73%) while Rotten Tomatoes has a mean of 47.2/100 (47.2%) and a median of 48/100 (48%). This discrepancy was surprising since we expected the rating systems to generally agree. 

### Comparison of Rating Systems

In order to further explore this unexpected discrepancy, we created a scatter plot comparing IMDb and Rotten Tomatoes. We also added the dimension of Age, which is the intended age group of each show. This allows us to visualize how shows for different age groups are rated. 

```{r results='markup'}

ggplot(data=tvdata, aes(x=IMDb, y=Rotten.Tomatoes, col=Age)) + geom_point() + ggtitle("Comparison of Rating Systems")+ scale_color_manual(breaks=c("7+","13+","16+","18+","all",NA),values=c("red", "blue", "green","orange","black","grey"))

print("Mean IMDb rating, 7+:")
mean(subset(tvdata,Age=="7+")$IMDb,na.rm = TRUE)
print("Mean RT rating, 7+:")
mean(subset(tvdata,Age=="7+")$Rotten.Tomatoes,na.rm = TRUE)
print("Mean IMDb rating, 13+:")
mean(subset(tvdata,Age=="13+")$IMDb,na.rm = TRUE)
print("Mean RT rating, 13+:")
mean(subset(tvdata,Age=="13+")$Rotten.Tomatoes,na.rm = TRUE)
print("Mean IMDb rating, 16+:")
mean(subset(tvdata,Age=="16+")$IMDb,na.rm = TRUE)
print("Mean RT rating, 16+:")
mean(subset(tvdata,Age=="16+")$Rotten.Tomatoes,na.rm = TRUE)
print("Mean IMDb rating, 18+:")
mean(subset(tvdata,Age=="18+")$IMDb,na.rm = TRUE)
print("Mean RT rating, 18+:")
mean(subset(tvdata,Age=="18+")$Rotten.Tomatoes,na.rm = TRUE)
print("Mean IMDb rating, all:")
mean(subset(tvdata,Age=="all")$IMDb,na.rm = TRUE)
print("Mean RT rating, all:")
mean(subset(tvdata,Age=="all")$Rotten.Tomatoes,na.rm = TRUE)
print("Mean IMDb rating, NA:")
mean(subset(tvdata,is.na(Age))$IMDb,na.rm = TRUE)
print("Mean RT rating, NA:")
mean(subset(tvdata,is.na(Age))$Rotten.Tomatoes,na.rm = TRUE)

```

This scatter plot further confirmed the fact that IMDb had higher ratings when compared to Rotten Tomatoes. There is a positive correlation between the two rating systems, but Rotten Tomatoes consistently has a lower overall rating. This helped us to form a new SMART question: what is the relationship between IMDb and Rotten Tomatoes?

This plot also gave us some information about how shows for different age groups are rated. Shows intended for 16+ had the highest overall rating, with an average of 7.3/10 on IMDb and 62.7/100 on Rotten Tomatoes. The 13+ age group at the lowest average IMDb score of 6.83/10 and shows intended for all ages had the lowest Rotten Tomatoes score with 47.7/100. It should be noted that shows with no intended age group listed (NA) had a lower Rotten Tomatoes score of 31.7/100, but since that represents a dearth of data concerning age group, the subset was neglected in future analysis.

## Streaming Platforms

### Ratings for Platforms

After comparing the rating systems to each other and then seeing how the intended age group affects ratings, we then turned to the different platforms. We compared the ratings for Disney, Hulu, Netflix, and Prime using boxplots.

```{r results='markup'}

netflixtv <- subset(tvdata, Netflix==1)
netflixtv$Platform <- "Netflix"
hulutv <- subset(tvdata, Hulu==1)
hulutv$Platform <- "Hulu"
primetv <- subset(tvdata, Prime.Video==1)
primetv$Platform <- "Prime"
disneytv <- subset(tvdata, Disney.==1)
disneytv$Platform <- "Disney"
plot.data <- rbind(netflixtv, hulutv, primetv, disneytv)

ggplot(plot.data, aes(x=Platform, y=IMDb, fill=Platform)) + geom_boxplot(outlier.shape = NA)+ggtitle("IMDb Rating for Platforms")+ylim(4,10)

ggplot(plot.data, aes(x=Platform, y=Rotten.Tomatoes, fill=Platform)) + geom_boxplot(outlier.shape = NA)+ggtitle("Rotten Tomatoes Rating for Platforms")

print("Mean IMDb rating, Disney:")
mean(disneytv$IMDb,na.rm=T)
print("Mean RT rating, Disney:")
mean(disneytv$Rotten.Tomatoes,na.rm=T)
print("Mean IMDb rating, Hulu:")
mean(hulutv$IMDb,na.rm=T)
print("Mean RT rating, Hulu:")
mean(hulutv$Rotten.Tomatoes,na.rm=T)
print("Mean IMDb rating, Netflix:")
mean(netflixtv$IMDb,na.rm=T)
print("Mean RT rating, Netflix:")
mean(netflixtv$Rotten.Tomatoes,na.rm=T)
print("Mean IMDb rating, Prime:")
mean(primetv$IMDb,na.rm=T)
print("Mean RT rating, Prime:")
mean(primetv$Rotten.Tomatoes,na.rm=T)

```

Much like how the rating systems had different overall distributions, they also painted different pictures for the average rating of shows on each platform. Using IMDb, Prime has the highest average rating of 7.15/10 and Disney has the lowest with 6.97/10. Looking at the Rotten Tomatoes ratings, however, Netflix has the highest average rating of 53.6/100, Hulu has the highest median score of 55/100, and Prime has the lowest average rating of 37.8/100. Prime is particularly interesting in this respect since it has the highest average rating with IMDb, but the lowest average rating with Rotten Tomatoes. This demonstrates that the question of "which platform has the highest average rating?" is not so straightforward.

### Age Groups by Platform

Next, we considered the relationship between platform and target age group. The frequency of target age groups differed significantly between different streaming platforms, thus making this a relevant feature to consider in later analysis.

```{r results='markup'}

ggplot(disneytv, aes(x=Age, fill=Age)) + geom_bar()+ggtitle("Targeted Age Groups (Disney)")+scale_x_discrete(limits=c("7+", "13+", "16+","18+","all"))

ggplot(hulutv, aes(x=Age, fill=Age)) + geom_bar()+ggtitle("Targeted Age Groups (Hulu)")+scale_x_discrete(limits=c("7+", "13+", "16+","18+","all"))

ggplot(netflixtv, aes(x=Age, fill=Age)) + geom_bar()+ggtitle("Targeted Age Groups (Netflix)")+scale_x_discrete(limits=c("7+", "13+", "16+","18+","all"))

ggplot(primetv, aes(x=Age, fill=Age)) + geom_bar()+ggtitle("Targeted Age Groups (Prime)")+scale_x_discrete(limits=c("7+", "13+", "16+","18+","all"))

nrow(subset(disneytv, Age=="all"))/nrow(disneytv)
nrow(subset(hulutv, Age=="16+"))/nrow(hulutv)
nrow(subset(netflixtv, Age=="18+"))/nrow(netflixtv)
nrow(subset(primetv, Age=="7+"))/nrow(primetv)

```

Starting with Disney, the most frequent targeted age group was all ages, with 36.8% of the Disney TV shows listed falling into that category. For Hulu, 16+ was the most common age group at 30.9%. The most common age group for Netflix was 18+, comprising 24.5% of its TV shows. Prime was more evenly distributed amongst age groups (apart from 13+, which was very low), but 7+ was the most common age group at 11.6%.

After seeing how much each streaming platform differed when it comes to age group, it would have been interesting to explore other demographics for the audience of each platform. Relevant datta in this respect could include gender, race, the number of views, and the actual age of the viewer (as opposed to the target age group). Since this data set did not include these features, this can be considered another limitation of the data. Our main focus was the TV show ratings, but we could have learned more about user preference and built a more detailed model with that additional information. 

```{r}

testfit1 <- lm(IMDb ~ Rotten.Tomatoes*Platform,data=plot.data)
summary(testfit1)
testfit2 <- lm(IMDb ~ Rotten.Tomatoes+Platform,data=plot.data)
summary(testfit2)
testfit3 <- lm(IMDb ~ Rotten.Tomatoes,data=plot.data)
summary(testfit3)

```


---------------------------
Test of Independence. To check whether IMDb and RRT is independent.


```{r }
contable1= table(netflixtv$Age, netflixtv$IMDb)
contable1
chitest1 = chisq.test(contable1)
chitest1

contable2 = table(netflixtv$Age, netflixtv$Rotten.Tomatoes)
contable2
chitest2 = chisq.test(contable2)
chitest2


```
Here p(0.02316)<0.05 rejects null Hypothesis. That's why IMDb and Rotten.Tomatoes for Netflix  are not independent.

```{r }
contable3= table(hulutv$Age, hulutv$IMDb)
contable3
chitest3 = chisq.test(contable3)
chitest3

contable4 = table(hulutv$Age, hulutv$Rotten.Tomatoes)
contable4
chitest4 = chisq.test(contable4)
chitest4

```
Here p<0.05 rejects null Hypothesis. That's why IMDb and Rotten.Tomatoes for Hulu  are not independent.

```{r}
contable5= table(primetv$Age, primetv$IMDb)
contable5
chitest5 = chisq.test(contable5)
chitest5

contable6 = table(primetv$Age, primetv$Rotten.Tomatoes)
contable6
chitest6 = chisq.test(contable6)
chitest6

```
Here p<0.05 rejects null Hypothesis. That's why IMDb and Rotten.Tomatoes for prime videos  are not independent.

```{r}
contable7= table(disneytv$Age, disneytv$IMDb)
contable7
chitest7 = chisq.test(contable7)
chitest7

contable8 = table(disneytv$Age, disneytv$Rotten.Tomatoes)
contable8
chitest8= chisq.test(contable8)
chitest8

```
Here p<0.05 rejects null Hypothesis. That's why IMDb and Rotten.Tomatoes for Disney  are not independent.

This step is to clear data, change some string data to number and some to factor.
```{r}
tvdata1 <- read.csv("tv_shows.csv", header=T, na.strings=c("","NA")) 
tvdata1$IMDb<-gsub("/10","",as.character(tvdata1$IMDb))
tvdata1$Rotten.Tomatoes<-gsub("/100","",as.character(tvdata1$Rotten.Tomatoes))
tvdata1$IMDb<-as.double(tvdata1$IMDb)
tvdata1$Rotten.Tomatoes<-as.double(tvdata1$Rotten.Tomatoes)

# total_plt <- tvdata$Netflix + tvdata$Hulu + tvdata$Prime.Video + tvdata$Disney.
# tvdata <- cbind(tvdata, total_plt)

tvdata1$Netflix = factor(tvdata1$Netflix)
tvdata1$Hulu = factor(tvdata1$Hulu)
tvdata1$Prime.Video = factor(tvdata1$Prime.Video)
tvdata1$Disney. = factor(tvdata1$Disney.)
tvdata1$Age = factor(tvdata1$Age)

str(tvdata1)
#View(tvdata)
```
To analyze the score relationships between two platforms, we need to drop the rows that have no IMDb scores and no Rotten.Tomatoes scores.
Meanwhile, the column of X, ID, Title and Type are useless, so we also drop these columns.
```{r}
tvdata_rank = subset(tvdata1, (!is.na(IMDb)) & (!is.na(Rotten.Tomatoes)))
tvdata_rank = subset(tvdata_rank, select = -c(X, ID, Title, Type))
#str(tvdata_rank)
# We've already done the above code 
hist(tvdata_rank$IMDb,breaks=20)
hist(tvdata_rank$Rotten.Tomatoes,breaks=20)
```

Make a pairs() plot with all the variables (quantitative and qualitative).
```{r}
loadPkg("lattice") # lattice and ggplot2 combined allow us to use the "pairs" function below 
pairs(tvdata_rank)
```

Make a corrplot() with only the numerical variables.
```{r}
loadPkg("corrplot")
tvdata_rank_num = subset(tvdata_rank, select = -c(Age, Netflix, Hulu, Prime.Video, Disney.))

cor_tvdata = cor(tvdata_rank_num)

corrplot(cor_tvdata, method = "number", type="upper")
# corrplot.mixed(cor_tvdata)
```

By using the variable Rotten.Tomatoes only, build a linear model with 1 independent variable to predict the IMDb.
```{r results='markup'}
model1 = lm(IMDb ~ Rotten.Tomatoes, data=tvdata_rank)
sum_md1 = summary(model1) # also for later use on inline codes
sum_md1

xkabledply(model1, title = paste("Model (num):", format(formula(model1)) ) )
```
From the results above, We can find there is a weak correlation between IMDb and Rotten.Tomatoes. And the correlation coefficient is 0.213.

Because there is only a weak correlation, I try to add the other variables into the model. Below is the result:
```{r}
model2 = lm(IMDb ~ ., data=tvdata_rank)
sum_md2 = summary(model2) # also for later use on inline codes
sum_md2
```
We can find that the adjusted correlation coefficient increases, but some variables are not significant, so try to drop it.


```{r}
table(tvdata_rank$Age)
```
The result is that for age13+, there are only 9 shows. Too little sample cause a large p-value. So we need to drop the factor of Age13+. We also drop Netflix, Prime.Video, Disney., because the p-values of these variables are not significant.
```{r}
tvdata_rank_no13 <- subset(tvdata_rank, Age!="13+")
tvdata_rank_no13 = subset(tvdata_rank_no13, select = -c(Netflix, Prime.Video, Disney.))
```

And then me make the third model as follow.
```{r}
model3 = lm(IMDb ~ ., data=tvdata_rank_no13)
sum_md3 = summary(model3) # also for later use on inline codes
sum_md3
```
We can find that all variables are significant and the adjusted r-squared is 0.246, much higher than the simple linear regression.

```{r results='markup'}
#vif_md3 = faraway::vif(model3)
#vif_md3

xkabledply(model3, title = paste("Model (num):", format(formula(model3)) ) )
xkablevif(model3)
```

```{r}
# aovresult <- anova(model1, model2, model3)
# aovresult
```
This is an Error because "models were not all fitted to the same size of dataset", so we redo the simple linear regression using tvdata_rank_no13 as model4.

```{r results='markup'}
model4 = lm(IMDb ~ Rotten.Tomatoes, data=tvdata_rank_no13)
sum_md4 = summary(model4) # also for later use on inline codes
sum_md4

xkabledply(model4, title = paste("Model (num):", format(formula(model4)) ) )
```

The concepts above can be extended naturally to models with interactions between numeric and factor variables.
```{r results='markup'}
model5 = lm(IMDb ~ . + Rotten.Tomatoes:Age, data=tvdata_rank_no13)
sum_md5 = summary(model5) # also for later use on inline codes
sum_md5

xkabledply(model5, title = paste("Model (num):", format(formula(model5)) ) )
```

Compared with these three models:
```{r results='markup'}
aovresult1 <- anova(model4, model3, model5)
aovresult1
xkabledply(aovresult1, title = "ANOVA comparison between the models")
```
Add one interaction, full model seems quite nice (although it's still a weak correlation).

IMDb = 23.464467 + 
-0.009186 * Year + 0.046692 * Rotten.Tomatoes + -0.209429 * Hulu
+ (-0.516497 + 0.000526 * Rotten.Tomatoes) * Age7+
+ (0.005813 + -0.007472 * Rotten.Tomatoes) * Age16+ 
+ (-0.453837 + -0.001010 * Rotten.Tomatoes) * Age18+ 
+ (0.609364 + -0.020359 * Rotten.Tomatoes) * Ageall


