---
title: "Group 1 Midterm"
author: "Evelyn Campo, Xiao Qi, Nusrat Prithee"
# date: "today"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r init, include=FALSE}
library(ezids)
library(ggplot2)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3)
```


```{r}
tvdata <- data.frame(read.csv("tv_shows.csv"))
str(tvdata)
```
This step is to clear data, change some string data to number and some to factor.
```{r}
tvdata$IMDb<-gsub("/10","",as.character(tvdata$IMDb))
tvdata$Rotten.Tomatoes<-gsub("/100","",as.character(tvdata$Rotten.Tomatoes))
tvdata$IMDb<-as.double(tvdata$IMDb)
tvdata$Rotten.Tomatoes<-as.double(tvdata$Rotten.Tomatoes)

# total_plt <- tvdata$Netflix + tvdata$Hulu + tvdata$Prime.Video + tvdata$Disney.
# tvdata <- cbind(tvdata, total_plt)

tvdata$Netflix = factor(tvdata$Netflix)
tvdata$Hulu = factor(tvdata$Hulu)
tvdata$Prime.Video = factor(tvdata$Prime.Video)
tvdata$Disney. = factor(tvdata$Disney.)
tvdata$Age = factor(tvdata$Age)

str(tvdata)
#View(tvdata)
```
To analyze the score relationships between two platforms, we need to drop the rows that have no IMDb scores and no Rotten.Tomatoes scores.
Meanwhile, the column of X, ID, Title and Type are useless, so we also drop these columns.
```{r}
tvdata_rank <- subset(tvdata, (!is.na(IMDb)) & (!is.na(Rotten.Tomatoes)))
tvdata_rank = subset(tvdata_rank, select = -c(X, ID, Title, Type))
str(tvdata_rank)

hist(tvdata_rank$IMDb,breaks=20)
hist(tvdata_rank$Rotten.Tomatoes,breaks=20)
```

Make a pairs() plot with all the variables (quantitative and qualitative).
```{r}
loadPkg("lattice") # lattice and ggplot2 combined allow us to use the "pairs" function below 
pairs(tvdata_rank)
```

Make a corrplot() with only the numerical variables.
```{r}
loadPkg("corrplot")
tvdata_rank_num = subset(tvdata_rank, select = -c(Age, Netflix, Hulu, Prime.Video, Disney.))

cor_tvdata = cor(tvdata_rank_num)

corrplot(cor_tvdata, method = "number", type="upper")
# corrplot.mixed(cor_tvdata)
```

By using the variable Rotten.Tomatoes only, build a linear model with 1 independent variable to predict the IMDb.
```{r}
model1 = lm(IMDb ~ Rotten.Tomatoes, data=tvdata_rank)
sum_md1 = summary(model1) # also for later use on inline codes
sum_md1
```
From the results above, We can find there is a weak correlation between IMDb and Rotten.Tomatoes. And the correlation coefficient is 0.213.

Because there is only a weak correlation, I try to add the other variables into the model. Below is the result:
```{r}
model2 = lm(IMDb ~ ., data=tvdata_rank)
sum_md2 = summary(model2) # also for later use on inline codes
sum_md2
```
We can find that the adjusted correlation coefficient increases, but some variables are not significant, so try to drop it.


```{r}
table(tvdata_rank$Age)
```
The result is that for age13+, there are only 9 shows. Too little sample cause a large p-value. So we need to drop the factor of Age13+. We also drop Netflix, Prime.Video, Disney., because the p-values of these variables are not significant.
```{r}
tvdata_rank_no13 <- subset(tvdata_rank, Age!="13+")
tvdata_rank_no13 = subset(tvdata_rank_no13, select = -c(Netflix, Prime.Video, Disney.))
```

And then me make the third model as follow.
```{r}
model3 = lm(IMDb ~ ., data=tvdata_rank_no13)
sum_md3 = summary(model3) # also for later use on inline codes
sum_md3
```
We can find that all variables are significant and the adjusted r-squared is 0.246, much higher than the simple linear regression.

```{r}
vif_md3 = faraway::vif(model3)
vif_md3
```

```{r}
# aovresult <- anova(model1, model2, model3)
# aovresult
```
This is an Error because "models were not all fitted to the same size of dataset", so we redo the simple linear regression using tvdata_rank_no13 as model4.

```{r}
model4 = lm(IMDb ~ Rotten.Tomatoes, data=tvdata_rank_no13)
sum_md4 = summary(model4) # also for later use on inline codes
sum_md4
```

The concepts above can be extended naturally to models with interactions between numeric and factor variables.
```{r}
model5 = lm(IMDb ~ . + Rotten.Tomatoes:Age, data=tvdata_rank_no13)
sum_md5 = summary(model5) # also for later use on inline codes
sum_md5
```

Compared with these three models:
```{r}
aovresult <- anova(model4, model3, model5)
aovresult
```
Add one interaction, full model seems quite nice (although it's still a weak correlation).

IMDb = 23.464467 + 
-0.009186 * Year + 0.046692 * Rotten.Tomatoes + -0.209429 * Hulu
+ (-0.516497 + 0.000526 * Rotten.Tomatoes) * Age7+
+ (0.005813 + -0.007472 * Rotten.Tomatoes) * Age16+ 
+ (-0.453837 + -0.001010 * Rotten.Tomatoes) * Age18+ 
+ (0.609364 + -0.020359 * Rotten.Tomatoes) * Ageall

