---
title: "Group 1 Midterm"
author: "Evelyn Campo, Xiao Qi, Nusrat Prithee, Roman"
date: "today"
# date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r init, include=FALSE}
library(ezids)
library(ggplot2)
library(dplyr)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
```

Importing the Dataset assigning "NA" to all blank cells
```{r}
tvdata <- read.csv("tv_shows.csv", header=T, na.strings=c("","NA")) 
```


Cleaning the Dataset
```{r}
tvdata$IMDb<-gsub("/10","",as.character(tvdata$IMDb)) #We had to delete the character /10 on every observation 
tvdata$Rotten.Tomatoes<-gsub("/100","",as.character(tvdata$Rotten.Tomatoes)) #We had to delete the character /100 on every observation 
tvdata$IMDb<-as.double(tvdata$IMDb) #as.double is a generic function. It is identical to as. numeric
tvdata$Rotten.Tomatoes<-as.double(tvdata$Rotten.Tomatoes) #as.double is a generic function. It is identical to as. numeric
tvdata = select(tvdata, -X, -ID, -Type) #This variables aren't of our interest

str(tvdata)
View(tvdata)
```

Counting missing values 
```{r}
sum(is.na(tvdata)) #Total NA
mean(is.na(tvdata))

#NA per variables
sum(is.na(tvdata$Title))
sum(is.na(tvdata$Year))
sum(is.na(tvdata$Age))
sum(is.na(tvdata$IMDb))
sum(is.na(tvdata$`Rotten Tomatoes`))
sum(is.na(tvdata$Netflix))
sum(is.na(tvdata$Hulu))
sum(is.na(tvdata$`Prime Video`))
sum(is.na(tvdata$`Disney+`))

```
Age has the most missing values in the Dataset, and IMDb is the other variable with NA values.

Streaming platforms into as.factor()
```{r}
tvdata$Netflix <- as.factor(tvdata$Netflix)  
tvdata$Hulu <- as.factor(tvdata$Hulu)  
tvdata$Prime.Video <- as.factor(tvdata$Prime.Video)
tvdata$Disney. <- as.factor(tvdata$Disney.)

```
The streaming platforms are categorical. 1 represents if the TV show is found on their platform otherwise 0.


Just undertanding a lilttle more about streaming platform variables and age
```{r}

print("Age:")
table(tvdata$Age)
print("Netflix:")
table(tvdata$Netflix)
print("Hulu:")
table(tvdata$Hulu)
print("Prime Video:")
table(tvdata$Prime.Video)
print("Disney:")
table(tvdata$Disney.)

```


What are the most targeted age groups for the TV shows by Netflix, Hulu, Prime, disney
Video? 

```{r}

print("Age:")
table(tvdata$Age)
prop.table(table(tvdata$Age))

ggplot(data = na.omit(tvdata), aes(Age)) + 
  geom_bar(col="red", 
                 fill="blue", 
                 alpha = .7) + # opacity
  labs(title="Most targeted age group for all platforms: 16+ ") +
  labs(x="Target age of viewers ", y="Frequency")

#I am taking this graph from above since it works for this question
```
people who are 16 and older are the most targeted age groups for the tv shows among the all steamming platfrom.

.Which year published the highest number of TV shows? 
```{r}
print("Year:")
table(tvdata$Year)
library(ggplot2)
ggplot(data=tvdata, aes(Year)) + 
  geom_histogram(bins = 30,
                 col="red", 
           fill="blue", 
           alpha = .7) + # opacity
  labs(title="Year during which the highest number of shows were published: 2017") +
  labs(x="Year tv show was published", y="Frequency")
#I am replacing the prevous code with the graphs from above to keep the same formatting
```
The highest number of TV shows were published in 2017 (685) and 2018 (562).And the histogram is right skewed which indicates that video publication is raising while time goes forward.

Normality check of Variables

```{r}

netflixtv <- subset(tvdata, Netflix==1)
hulutv <- subset(tvdata, Hulu==1)
primetv <- subset(tvdata, Prime.Video==1)
disneytv <- subset(tvdata, Disney.==1)

```

Which streaming platform has the highest average rating (according to Rotten 
Tomatoes and IMDb)? 

```{r q06}
net_rating<-t.test(netflixtv$IMDb,netflixtv$Rotten.Tomatoes)
net_rating
hulu_rating<-t.test(hulutv$IMDb,hulutv$Rotten.Tomatoes)
hulu_rating
prime_rating<-t.test(primetv$IMDb,primetv$Rotten.Tomatoes)
prime_rating
disney_rating<-t.test(disneytv$IMDb,disneytv$Rotten.Tomatoes)
disney_rating
```
 Prime Videos has the highest average Imdb rating  which is 7.152538. Also, Netflix has the highest average RTT rating  which is 53.559107
 
Normality check for the variable IMDb and Rotten.Tomatoes
```{r}
shapiro.test(netflixtv$IMDb)
shapiro.test(netflixtv$Rotten.Tomatoes)

library(ggplot2)
ggplot(data=netflixtv,aes(x=IMDb,)) + 
geom_histogram(col="red", 
               fill="green", 
               alpha = .7) + 
labs(title="Netflix IMDb histogram") +
labs(x="IMDB", y="frequency")

ggplot(data=netflixtv,aes(x=Rotten.Tomatoes,)) + 
geom_histogram(col="red", 
               fill="green", 
               alpha = .7) + 
labs(title="Netflix Rotten.Tomatoes histogram") +
labs(x="Rotten.Tomatoes", y="frequency")


```
Here p value for IMDb and Rotten.Tomatoes is less than 0.05 for Netflix. That's mean these are not normally distributed. Histogram of IMDb is right-skewed also Histogram of Rotten.Tomatoes is slightly left-skewed.
```{r}
shapiro.test(hulutv$IMDb)
shapiro.test(hulutv$Rotten.Tomatoes)

library(ggplot2)
ggplot(data=hulutv,aes(x=IMDb,)) + 
geom_histogram(col="red", 
               fill="green", 
               alpha = .7) + 
labs(title="Hulu IMDb histogram") +
labs(x="IMDB", y="frequency")

ggplot(data=hulutv,aes(x=Rotten.Tomatoes,)) + 
geom_histogram(col="red", 
               fill="green", 
               alpha = .7) + 
labs(title="hulu Rotten.Tomatoes histogram") +
labs(x="Rotten.Tomatoes", y="frequency")

```
Here p value for IMDb and Rotten.Tomatoes is less than 0.05 for Hulu. That's mean these are not normally distributed. Histogram of IMDb is right-skewed also Histogram of Rotten.Tomatoes is slightly Bimodal.
```{r}
shapiro.test(primetv$IMDb)
shapiro.test(primetv$Rotten.Tomatoes)
library(ggplot2)
ggplot(data=primetv,aes(x=IMDb,)) + 
geom_histogram(col="red", 
               fill="green", 
               alpha = .7) + 
labs(title="prime IMDb histogram") +
labs(x="IMDB", y="frequency")

ggplot(data=primetv,aes(x=Rotten.Tomatoes,)) + 
geom_histogram(col="red", 
               fill="green", 
               alpha = .7) + 
labs(title="prime Rotten.Tomatoes histogram") +
labs(x="Rotten.Tomatoes", y="frequency")

```
Here p value for IMDb and Rotten.Tomatoes is less than 0.05 for Prime Videos. That's mean these are not normally distributed. Histogram of IMDb is right-skewed also Histogram of Rotten.Tomatoes is Randomly distribution.

```{r}
shapiro.test(disneytv$IMDb)
shapiro.test(disneytv$Rotten.Tomatoes)
library(ggplot2)
ggplot(data=disneytv,aes(x=IMDb,)) + 
geom_histogram(col="red", 
               fill="green", 
               alpha = .7) + 
labs(title="disney IMDb histogram") +
labs(x="IMDB", y="frequency")

ggplot(data=disneytv,aes(x=Rotten.Tomatoes,)) + 
geom_histogram(col="red", 
               fill="green", 
               alpha = .7) + 
labs(title="disney Rotten.Tomatoes histogram") +
labs(x="Rotten.Tomatoes", y="frequency")

```
Here p value for IMDb and Rotten.Tomatoes is less than 0.05 for Disney+. That's mean these are not normally distributed. Histogram of IMDb is slightly right-skewed and randomly distributed. also Histogram of Rotten.Tomatoes is slightly left-skewed and randomly distribution.
----------------
```{r}
hist(tvdata$IMDb,breaks=20)
hist(tvdata$Rotten.Tomatoes,breaks=20)

```

```{r}

ggplot(data=tvdata, aes(x=IMDb, y=Rotten.Tomatoes, col=Age)) + geom_point() + ggtitle("Comparison of Rating Systems")

```


```{r}

netflixtv <- subset(tvdata, Netflix==1)
netflixtv$Platform <- "Netflix"
hulutv <- subset(tvdata, Hulu==1)
hulutv$Platform <- "Hulu"
primetv <- subset(tvdata, Prime.Video==1)
primetv$Platform <- "Prime"
disneytv <- subset(tvdata, Disney.==1)
disneytv$Platform <- "Disney"
plot.data <- rbind(netflixtv, hulutv, primetv, disneytv)
ggplot(plot.data, aes(x=Platform, y=IMDb)) + geom_boxplot(outlier.shape = NA)+
  scale_y_continuous(limits = (3:10))
ggplot(plot.data, aes(x=Platform, y=Rotten.Tomatoes)) + geom_boxplot(outlier.shape = NA)

```


```{r}

testfit1 <- lm(IMDb ~ Rotten.Tomatoes*Platform,data=plot.data)
summary(testfit1)
testfit2 <- lm(IMDb ~ Rotten.Tomatoes+Platform,data=plot.data)
summary(testfit2)
testfit3 <- lm(IMDb ~ Rotten.Tomatoes,data=plot.data)
summary(testfit3)

```
---------------------------
Test of Independence. To check whether IMDb and RRT is independent.
```{r }
contable1= table(netflixtv$Age, netflixtv$IMDb)
contable1
chitest1 = chisq.test(contable1)
chitest1

contable2 = table(netflixtv$Age, netflixtv$Rotten.Tomatoes)
contable2
chitest2 = chisq.test(contable2)
chitest2


```
Here p(0.02316)<0.05 rejects null Hypothesis. That's why IMDb and Rotten.Tomatoes for Netflix  are not independent.

```{r }
contable3= table(hulutv$Age, hulutv$IMDb)
contable3
chitest3 = chisq.test(contable3)
chitest3

contable4 = table(hulutv$Age, hulutv$Rotten.Tomatoes)
contable4
chitest4 = chisq.test(contable4)
chitest4

```
Here p<0.05 rejects null Hypothesis. That's why IMDb and Rotten.Tomatoes for Hulu  are not independent.

```{r}
contable5= table(primetv$Age, primetv$IMDb)
contable5
chitest5 = chisq.test(contable5)
chitest5

contable6 = table(primetv$Age, primetv$Rotten.Tomatoes)
contable6
chitest6 = chisq.test(contable6)
chitest6

```
Here p<0.05 rejects null Hypothesis. That's why IMDb and Rotten.Tomatoes for prime videos  are not independent.

```{r}
contable7= table(disneytv$Age, disneytv$IMDb)
contable7
chitest7 = chisq.test(contable7)
chitest7

contable8 = table(disneytv$Age, disneytv$Rotten.Tomatoes)
contable8
chitest8= chisq.test(contable8)
chitest8

```
Here p<0.05 rejects null Hypothesis. That's why IMDb and Rotten.Tomatoes for Disney  are not independent.

This step is to clear data, change some string data to number and some to factor.
```{r}
tvdata1 <- read.csv("tv_shows.csv", header=T, na.strings=c("","NA")) 
tvdata1$IMDb<-gsub("/10","",as.character(tvdata1$IMDb))
tvdata1$Rotten.Tomatoes<-gsub("/100","",as.character(tvdata1$Rotten.Tomatoes))
tvdata1$IMDb<-as.double(tvdata1$IMDb)
tvdata1$Rotten.Tomatoes<-as.double(tvdata1$Rotten.Tomatoes)

# total_plt <- tvdata$Netflix + tvdata$Hulu + tvdata$Prime.Video + tvdata$Disney.
# tvdata <- cbind(tvdata, total_plt)

tvdata1$Netflix = factor(tvdata1$Netflix)
tvdata1$Hulu = factor(tvdata1$Hulu)
tvdata1$Prime.Video = factor(tvdata1$Prime.Video)
tvdata1$Disney. = factor(tvdata1$Disney.)
tvdata1$Age = factor(tvdata1$Age)

str(tvdata1)
#View(tvdata)
```
To analyze the score relationships between two platforms, we need to drop the rows that have no IMDb scores and no Rotten.Tomatoes scores.
Meanwhile, the column of X, ID, Title and Type are useless, so we also drop these columns.
```{r}
tvdata_rank = subset(tvdata1, (!is.na(IMDb)) & (!is.na(Rotten.Tomatoes)))
tvdata_rank = subset(tvdata_rank, select = -c(X, ID, Title, Type))
#str(tvdata_rank)
# We've already done the above code 
hist(tvdata_rank$IMDb,breaks=20)
hist(tvdata_rank$Rotten.Tomatoes,breaks=20)
```

Make a pairs() plot with all the variables (quantitative and qualitative).
```{r}
loadPkg("lattice") # lattice and ggplot2 combined allow us to use the "pairs" function below 
pairs(tvdata_rank)
```

Make a corrplot() with only the numerical variables.
```{r}
loadPkg("corrplot")
tvdata_rank_num = subset(tvdata_rank, select = -c(Age, Netflix, Hulu, Prime.Video, Disney.))

cor_tvdata = cor(tvdata_rank_num)

corrplot(cor_tvdata, method = "number", type="upper")
# corrplot.mixed(cor_tvdata)
```

By using the variable Rotten.Tomatoes only, build a linear model with 1 independent variable to predict the IMDb.
```{r results='markup'}
model1 = lm(IMDb ~ Rotten.Tomatoes, data=tvdata_rank)
sum_md1 = summary(model1) # also for later use on inline codes
sum_md1

xkabledply(model1, title = paste("Model (num):", format(formula(model1)) ) )
```
From the results above, We can find there is a weak correlation between IMDb and Rotten.Tomatoes. And the correlation coefficient is 0.213.

Because there is only a weak correlation, I try to add the other variables into the model. Below is the result:
```{r}
model2 = lm(IMDb ~ ., data=tvdata_rank)
sum_md2 = summary(model2) # also for later use on inline codes
sum_md2
```
We can find that the adjusted correlation coefficient increases, but some variables are not significant, so try to drop it.


```{r}
table(tvdata_rank$Age)
```
The result is that for age13+, there are only 9 shows. Too little sample cause a large p-value. So we need to drop the factor of Age13+. We also drop Netflix, Prime.Video, Disney., because the p-values of these variables are not significant.
```{r}
tvdata_rank_no13 <- subset(tvdata_rank, Age!="13+")
tvdata_rank_no13 = subset(tvdata_rank_no13, select = -c(Netflix, Prime.Video, Disney.))
```

And then me make the third model as follow.
```{r}
model3 = lm(IMDb ~ ., data=tvdata_rank_no13)
sum_md3 = summary(model3) # also for later use on inline codes
sum_md3
```
We can find that all variables are significant and the adjusted r-squared is 0.246, much higher than the simple linear regression.

```{r results='markup'}
#vif_md3 = faraway::vif(model3)
#vif_md3

xkabledply(model3, title = paste("Model (num):", format(formula(model3)) ) )
xkablevif(model3)
```

```{r}
# aovresult <- anova(model1, model2, model3)
# aovresult
```
This is an Error because "models were not all fitted to the same size of dataset", so we redo the simple linear regression using tvdata_rank_no13 as model4.

```{r results='markup'}
model4 = lm(IMDb ~ Rotten.Tomatoes, data=tvdata_rank_no13)
sum_md4 = summary(model4) # also for later use on inline codes
sum_md4

xkabledply(model4, title = paste("Model (num):", format(formula(model4)) ) )
```

The concepts above can be extended naturally to models with interactions between numeric and factor variables.
```{r results='markup'}
model5 = lm(IMDb ~ . + Rotten.Tomatoes:Age, data=tvdata_rank_no13)
sum_md5 = summary(model5) # also for later use on inline codes
sum_md5

xkabledply(model5, title = paste("Model (num):", format(formula(model5)) ) )
```

Compared with these three models:
```{r results='markup'}
aovresult1 <- anova(model4, model3, model5)
aovresult1
xkabledply(aovresult1, title = "ANOVA comparison between the models")
```
Add one interaction, full model seems quite nice (although it's still a weak correlation).

IMDb = 23.464467 + 
-0.009186 * Year + 0.046692 * Rotten.Tomatoes + -0.209429 * Hulu
+ (-0.516497 + 0.000526 * Rotten.Tomatoes) * Age7+
+ (0.005813 + -0.007472 * Rotten.Tomatoes) * Age16+ 
+ (-0.453837 + -0.001010 * Rotten.Tomatoes) * Age18+ 
+ (0.609364 + -0.020359 * Rotten.Tomatoes) * Ageall


